---
title: "Practical Machine Learning Project"
author: "trungnt"
date: "April 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement-a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset)

```{r, echo=FALSE}
setwd("/home/trungnt/Documents/Learn/ML/Coursera/DS-JHU/C8-Practical ML/project")
```

## Data Processing

1. Load the required packages

```{r, message=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
```  

2. Loading data set
```{r}
trainDF <- read.csv("data/pml-training.csv", na.strings = c("NA", ""))
testDF  <- read.csv("data/pml-testing.csv",  na.strings = c("NA", ""))
str(trainDF)
```  
The training dataset has 19622 observations and 160 variables, and the testing data set contains 20 observations and the same variables as the training set. We are trying to predict the outcome of the variable classe in the training set.  

3. Data cleaning  

###### Delete columns that contain any missing values
```{r}
trainDF <- trainDF[, colSums(is.na(trainDF)) == 0]
testDF  <- testDF[, colSums(is.na(testDF)) == 0]
```   

###### Delete columns have little predicting power for the outcome $classe$
```{r}
trainDF <- trainDF[, -c(1:7)]
testDF  <- testDF[, -c(1:7)]
dim(trainDF)
dim(testDF)
```  
Final, we have train set with 53 columns and 19622 rows, test set has same columns and 20 rows. 

4. Split train set  
We nead split train set into a training set (train, 70%) for prediction and a validation set (valid 30%) to compute the out-of-sample errors.
```{r}
set.seed(10) 
indTrain <- createDataPartition(trainDF$classe, p = 0.7, list = FALSE)
train <- trainDF[indTrain,  ]
valid <- trainDF[-indTrain, ]
```  

### Classification with Random forest algorithms  
1. Training  
```{r}
control <- trainControl(method = "cv", number = 5)
model_rf <- train(classe ~ ., data = train, method = "rf", trControl = control)
print(model_rf, digits = 4)
```  

```{r}
plot(model_rf, main="Accuracy of Random forest model by number of predictors")
```  

2. Testing
Testing outcomes using validation set
```{r}
predict_rf <- predict(model_rf, valid)
(conf_rf <- confusionMatrix(valid$classe, predict_rf))
```  

For this dataset, random forest method is way better than classification tree method. The accuracy rate is 0.991, and so the out-of-sample error rate is 0.009  

3. Prediction  
Use $random forests$ to predict the outcome variable classe for the testing set.  
```{r}
(predict(model_rf, testDF))
```